---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 320
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version-ago/GenerateIndexPedigree)](https://CRAN.R-project.org/package=GenerateIndexPedigree)
[![CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/GenerateIndexPedigree)](https://CRAN.R-project.org/package=GenerateIndexPedigree)
<!-- badges: end -->

# GenerateIndexPedigree

A series of functions to generate a pedigree around a given person with a known genetic mutation, based on US census data. Pedigree will also track mutation status of all members. Pedigree is built down to children, back to grandparents, and out to cousins. There is also a function to generate pedigrees without genetic mutations.

## Installation

You can install the package directly from CRAN with:

```{r CRAN_installation, eval=FALSE}
install.packages("GenerateIndexPedigree")
```

Or you can install the development, GitHub version of this package with:

```{r GitHub_installation, eval=FALSE}
install.packages("remotes")
remotes::install_github("katelynqueen98/GenerateIndexPedigree")
```

## Examples

The main functions of the package, `GenerateIndexPedigree()` and `GeneratePedigree()` provide many parameter options. Here is a basic pedigree of a 30-year-old female with a mutation in BRCA1, which has a *de novo* mutation rate of 0.005.

```{r default_example}
library(GenerateIndexPedigree)
set.seed(123)
GenerateIndexPedigree(age = 30,
                      sex = "F",
                      variant = "BRCA1",
                      denovo_rate = 0.005)
```

By default, all children that will ever be born from an individual are included, regardless of if they have been born yet. This is shown by negative ages, indicating the child will be born in a future year. These individuals can be excluded by setting parameter `keep_future_children` to `FALSE`

```{r no_future_children_example}
set.seed(123)
GenerateIndexPedigree(age = 30,
                      sex = "F",
                      variant = "BRCA1",
                      denovo_rate = 0.005,
                      keep_future_children = FALSE)

```

Additionally, the *de novo* status of each generational mutation (index, parent, grandparent) is simulated using either the default rate of 2e-06 or the rate given in parameter `denovo_rate`. If the *de novo* status of the index case is known, it can be supplied in parameter `is_denovo`.

```{r denovo_example}
set.seed(123)
GenerateIndexPedigree(age = 30,
                      sex = "F",
                      variant = "BRCA1",
                      keep_future_children = FALSE,
                      is_denovo = TRUE,
                      verbose = TRUE)

```

Finally, a pedigree without any mutation information can be generated using the `GeneratePedigree()` function.

```{r no_mutation_example}
set.seed(123)
GeneratePedigree(age = 30,
                 sex = "F",
                 keep_future_children = FALSE)
```


## Simulating many pedigrees

Average family sizes can be recapitulated by simulating large numbers of pedigrees. Here, we simulate 10,000 pedigrees from 30-year-olds with a BRCA1 mutation. We can also track the number of positive cascade tests generated in family members of testing age, 20 - 65.

```{r generations_simulation}

# simulation - 10K pedigrees of given age, random sex
PedigreeSimulation <- function(indexAge,
                               nsim = 10000,
                               verbose = FALSE) {
  # set up 
  pedigreeSummary <- data.frame(indexAge = numeric(nsim),
                                indexVariant = character(nsim),
                                indexSex = character(nsim),
                                DNM = logical(nsim),
                                numKids = numeric(nsim),
                                possibleNumCascadeTests = numeric(nsim),
                                numCascadeTests = numeric(nsim),
                                numCascadePositives = numeric(nsim))
  
  # simulate
  for (i in 1:nsim) {
    variant <- "BRCA1"
    
    # simulate pedigree
    sex <- sample(c("M", "F"), 1)
    ped <- GenerateIndexPedigree(age = indexAge,
                                 sex = sex,
                                 variant = variant,
                                 denovo_rate = 0.005)
    
    # fix data type
    ped$Age <- as.numeric(ped$Age)
    
    # summary stats - index
    pedigreeSummary$indexAge[i]     <- indexAge
    pedigreeSummary$indexVariant[i] <- variant
    pedigreeSummary$indexSex[i]     <- sex
    pedigreeSummary$numKids[i]      <- sum(grepl("Child", ped$RelationshipToIndex))
    pedigreeSummary$DNM[i]          <- ifelse(isTRUE(ped$DeNovo[which(ped$RelationshipToIndex == "Self")]), TRUE, FALSE)
    
    # summary stats - pedigree
    ped_inAgeRange <- ped[which(ped$Age >= (indexAge - 10) & 
                                  ped$Age <= (indexAge + 35)), ]
    pedigreeSummary$possibleNumCascadeTests[i] <- nrow(ped_inAgeRange) - 1
    pedigreeSummary$numCascadeTests[i] <- max(nrow(ped_inAgeRange[which(ped_inAgeRange$UptakeTesting == "Yes"), ]) - 1, 0)
    pedigreeSummary$numCascadePositives[i] <- max(nrow(ped_inAgeRange[which(ped_inAgeRange$UptakeTesting == "Yes" 
                                            & ped_inAgeRange$Variant == variant), ]) - 1, 0)
  }
  
  # fix class of de novo
  pedigreeSummary$DNM <- as.logical(pedigreeSummary$DNM)
  
  # to return...
  pedigreeSummary
  print(paste0("**************** ", format(nsim, big.mark = ",", scientific = FALSE),
               " pedigrees of 30 year old index case born in ", 
               2026 - indexAge, " ****************"))
  print("Number of Children of Index:")
  print(summary(pedigreeSummary$numKids))
  print("Number of Relatives of testable age:")
  print(summary(pedigreeSummary$possibleNumCascadeTests))
  print("Number of Cascade Tests:")
  print(summary(pedigreeSummary$numCascadeTests))
  print("Number of Cascade Positives:")
  print(summary(pedigreeSummary$numCascadePositives))
  print("Index case de novo mutation rate:")
  print(round(sum(pedigreeSummary$DNM)/nsim, 4))
}

PedigreeSimulation(30)

```

